---
title: "Introduction to Suitability Mapping"
date: "Week 8: 29/11/2021"
---

<style type: "text/css">
h1.title {
	font-size: 20px;
}

h4.date {
	font-size: 18px;
}

h1 {
	font-size: 18px;
}

h2 {
	font-size: 18px;
}

body{/* Normal */
	font-size: 13px;
	text-align: justify;
}

p.comment {
background-color: #DBDBDB;
padding: 10px;
border: 1px solid black;
border-radius: 5px;
}

code.r{
  font-size: 10px;
}
pre {
  font-size: 12px
}
</style>

<hr style="border:2px solid gray"> </hr>

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = "/Users/anwarmusah/Documents/GITHUB/UCLPG-MSC-SGDS/Dataset/T1_WK8")
knitr::opts_chunk$set(cache = TRUE)
```

## **1.1. Let's begin**
**Background**: Risk models derived from environmental, sociodemographic and topological factors are becoming increasingly available for open research. Such models have broadly shown to be useful in delineating geographical areas of risk or suitability for a certain outcomes. So far, we have covered different ways for modelling the spatial distribution of outcomes and estimating its occurrence or risk, we have explored this through areal-based, geostatistical and nonparametric methods. Today, we are going to explore this from a different lens using a knowledge-driven approach. 

**OBJECTIVES**: To provide an introductory overview to the applicability of knowledge-driven methods, in particular, we are going to learn the **Multi-Criteria Decision Approach** (MCDA) which is a method that use decision rules from existing knowledge to identify areas potential suitability for an outcome. It is especially useful in data-sparse situations, or when for the first time exploring the potential geographical limits of certain outcome. 
For instance, using modest number of raster layers such as population density, urbanisation, approximating to street segments, house prices and deprivation; it is possible to use such information in determining regions for which crime events such as burglaries are likely to occur, or suitable in that matter. This approach has been widely used in a number of disciplines over the last decades, and has gained prominence in public health related fields such as vector-borne disease prevention, and disaster sciences such as landslides. We will learn how to apply these methods to the two context.  

</br>

## **1.2. Datasets & setting up the work directory** 
Before you begin do make sure to download all data from [**HERE**](https://github.com/UCLPG-MSC-SGDS/GEOG0114-PSA-WK8/raw/main/Datasets%20for%20Kenya.zip). If you are working from 
a UCL workstation, do create a folder called "**Week 8**" within your "**GEOG0114**" folder stored in the N-drive.

Extract all data from the zip folder and stored into "**Week 8**" folder. Open a new R script and set the work directory 
to **Week 8's** folder (i.e., **Home (N:) > GEOG0114 > Week 8**)

```{r, eval = FALSE}
# Set working directory to Week 6 folder
setwd("N:/GEOG0114/Week 8")
```

</br>

## **1.3. Loading and installing packages**


<hr style="border:2px solid gray"> </hr>
